---
title: "Survival Analysis Project"
author: "Valentin CAPIEU"
date: "2025-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load libraries
library(tidyverse)
library(data.table)

```
```{r}
# Load dataset
data <- fread("data/brca_metabric_clinical_data.tsv")
```

```{r}
# Check if Sample ID and Patient ID are identical
identical_ids <- identical(data$`Sample ID`, data$`Patient ID`)
cat("Are 'Sample ID' and 'Patient ID' identical? →", identical_ids, "\n")

# Check unique values in cancer type, Study ID, Sample Type, Number of Samples Per Patient
unique_cancer_types <- unique(data$`Cancer Type`)
unique_study_ids <- unique(data$`Study ID`)
unique_sample_types <- unique(data$`Sample Type`)
unique_numbers_of_samples_per_patient<-unique(data$`Number of Samples Per Patient`)

#cat("Unique values in 'Cancer Type':", length(unique_cancer_types), "\n")
#cat("Unique values in 'Study ID':", length(unique_study_ids), "\n")
cat("Unique values in 'Study ID','Cancer Type','Sample Type','Number of Samples Per Patient' : ", length(unique_cancer_types), length(unique_study_ids), length(unique_sample_types), length(unique_numbers_of_samples_per_patient), "\n")

# If assumptions are true, remove redundant columns
if (identical_ids) {
  data <- subset(data, select = -`Sample ID`)
}

if (length(unique_cancer_types) == 1) {
  data <- subset(data, select = -`Cancer Type`)
}

if (length(unique_study_ids) == 1) {
  data <- subset(data, select = -`Study ID`)
}

if (length(unique_sample_types) == 1) {
  data <- subset(data, select = -`Sample Type`)
}

if (length(unique_numbers_of_samples_per_patient) == 1) {
  data <- subset(data, select = -`Number of Samples Per Patient`)
}

# Confirm structure after cleaning
cat("\nColumns remaining after cleaning:\n")
print(names(data))
```
```


```{r}
str(data)

# Separate continuous and categorical covariates
continuous_vars <- names(data)[sapply(data, is.numeric)]
categorical_vars <- names(data)[sapply(data, function(x) is.character(x) | is.factor(x))]

```

```{r}
# Create a folder for plots
dir.create("plots", showWarnings = FALSE)

for (v in continuous_vars) {
  p <- ggplot(data, aes_string(x = v)) +
    geom_histogram(bins = 30, color = "black", fill = "skyblue") +
    theme_minimal() +
    labs(title = paste("Histogram of", v))
  
  # ggsave(filename = paste0("plots/hist_", v, ".png"), plot = p, width = 6, height = 4
  p
}

```

```{r}
for (v in categorical_vars) {
  p <- ggplot(data, aes_string(x = v)) +
    geom_bar(fill = "orange") +
    theme_minimal() +
    labs(title = paste("Bar plot of", v), x = v, y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  ggsave(filename = paste0("plots/bar_", v, ".png"), plot = p, width = 6, height = 4)
}

```

